---
title: "World Development Indicators: Key Insights"
author: "Alan"
date: 2025-06-06
format: 
  revealjs:
    theme: serif
    transition: fade
    slide-number: true
    toc: true
editor: visual
bibliography: references.bib
---

```{r}
#| label: setup
#| echo: false

library(ggplot2)
library(dplyr)
library(readr)
library(knitr)

wdi <- read_csv("wdi_2022.csv")
```

## Welcome

Welcome to this presentation on the 2022 World Development Indicators.  
We’ll explore GDP, life expectancy, education, and population trends.

---

## GDP vs Life Expectancy

- Wealthier countries often enjoy longer life expectancies.
- Higher GDP typically correlates with better public health outcomes.

```{r}
#| label: fig-gdp-life-slide
#| echo: false
#| fig-cap: "GDP vs Life Expectancy (2022)"

ggplot(wdi, aes(x = gdp_per_capita, y = life_expectancy)) +
  geom_point(color = "steelblue") +
  labs(x = "GDP per Capita (USD)", y = "Life Expectancy (Years)") +
  theme_minimal()
```

---

## Top 10 Most Populous Countries

- Population is highly concentrated in a few nations.
- Useful for understanding market size, healthcare needs, etc.

```{r}
#| label: fig-population
#| echo: false
#| fig-cap: "Top 10 Most Populous Countries (2022)"

wdi %>%
  filter(!is.na(total_population)) %>%
  arrange(desc(total_population)) %>%
  slice_head(n = 10) %>%
  ggplot(aes(x = reorder(country, total_population), y = total_population)) +
  geom_bar(stat = "identity", fill = "darkgreen") +
  coord_flip() +
  labs(x = "Country", y = "Total Population") +
  theme_minimal()
```

---

## Summary Statistics

Key statistics across selected indicators.

```{r}
#| label: tbl-summary-slide
#| echo: false
#| tbl-cap: "Summary Statistics of Development Indicators (2022)"

summary_table <- wdi %>%
  summarise(
    `Mean GDP per Capita` = mean(gdp_per_capita, na.rm = TRUE),
    `Median GDP per Capita` = median(gdp_per_capita, na.rm = TRUE),
    `Min GDP per Capita` = min(gdp_per_capita, na.rm = TRUE),
    `Max GDP per Capita` = max(gdp_per_capita, na.rm = TRUE),

    `Mean Life Expectancy` = mean(life_expectancy, na.rm = TRUE),
    `Median Life Expectancy` = median(life_expectancy, na.rm = TRUE),
    `Min Life Expectancy` = min(life_expectancy, na.rm = TRUE),
    `Max Life Expectancy` = max(life_expectancy, na.rm = TRUE),

    `Mean Unemployment Rate` = mean(unemployment_rate, na.rm = TRUE),
    `Mean Education % of GDP` = mean(education_expenditure_gdp_share, na.rm = TRUE),
    `Mean Total Population` = mean(total_population, na.rm = TRUE)
  ) %>%
  t() %>%
  as.data.frame()

colnames(summary_table) <- "Value"
knitr::kable(summary_table)
```

---

## Conclusion

- Countries with higher GDP tend to have longer life expectancy.
- Population and education investment vary widely across countries.
- Insightful indicators help guide economic and public policy.

---
